The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2017-02-15 manuscript/manuscript.Rnw
Indicate genes assoc. with adverse pregnancy
Show which genes are associated with adverse pregnancy outcomes using
literature search from pubmed and link to those results in Table 1.

@@

placenta.lineage.specific.table <- function(analysis,output=TRUE,lineage.table=placenta.lineage.specific.analysis,members=5,orderbyfc=FALSE,fdr.cutoff=1) {
    temp.table <- 
        lineage.table[analysis.name==analysis &
                      p.value <= 0.05 &
                      fdr <= fdr.cutoff][order(p.value),
                                       list(human_name,p.value,fdr,fc),]
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[`FDR`<=fdr.cutoff,][order(`p (t-test)`),]
        if (orderbyfc) {
            display.table <- temp.table[order(-abs(`Fold Change`)),]
        }
        display.table <- display.table[1:members,]

        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        display.table[,`Human Name`:=NULL]
        # setnames(display.table,"Human Symbol","\\multicolumn{1}{p{0.5in}}{Human Symbol}")6
        setnames(display.table,"Fold Change","{$\\text{log}_2$ FC}")
        setnames(display.table,"FDR","{FDR}")
        setnames(display.table,"p (t-test)","{p (t-test)}")
        print(xtable(display.table,
                     display=c("s","s","e","e","f"),
                     align=c("c","p{0.5in}","S[table-format=1.2e2]","S[table-format=1.2e2]","S[table-format=2.2]")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=FALSE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

@@

\begin{table}[tbh]
  \footnotesize
<<human.lineage.specific,echo=FALSE,results="asis">>=
hum.all.lin.spec <-
    placenta.lineage.specific.table("hum.all",
                                    lineage.table=human.mouse.lineage.specific.analysis,
                                    members=40,
                                    orderbyfc=TRUE,
                                    fdr.cutoff=0.001
                                    )
setkey(hum.all.lin.spec,"Human Symbol")
@ 
\caption{\lineagetableexplanationtypeiiishort{\human}{all other analyzed species}}
\label{tab:human_lineage_specific}
\end{table}

>>>>

To identify genes which have evidence of association with adverse
pregnancy outcomes and other placenta-relevant terms of interest, we
used the “pubmed_to_grid” component of
Function2Gene~\cite{Armstrong2008:function2gene} which conducts a term
by gene search of PubMed results.

@@

literature.mining <-
    data.table(read.table("../literature_mining/human_lineage_vs_terms.txt",
                          sep="\t",skip=0,stringsAsFactors=FALSE,
                          check.names=FALSE,header=TRUE),
               keep.rownames=TRUE)
setnames(literature.mining,"rn","human_name")
setkey(literature.mining,"human_name")
literature.mining[,`Adverse Pregnancy Outcomes`:=`preeclampsia`+
                       IUGR+`placenta previa`+`gestational diabetes`+
                       `premature birth`,
                  by=human_name]
literature.mining.terms <-
    paste0("%28%22preeclampsia%22[All%20Fields]%20OR%20",
           "%22pre-eclampsia%22[All%20Fields]%20OR%20",
           "%22eclampsia%22[All%20Fields]%20OR%20",
           "%22IUGR%22[All%20Fields]%20OR%20",
           "%22iugr%22[All%20Fields]%20OR%20",
           "%22intrauterine%20growth%20restriction%22[All%20Fields]%20OR%20",
           "%22placenta%20previa%22[All%20Fields]%20OR%20",
           "%22gestational%20diabetes%22[All%20Fields]%20OR%20",
           "%22premature%20birth%22[All%20Fields]%20OR%20",
           "%22premature%20labor%22[All%20Fields]%29")
placenta.lineage.specific.table <- function(analysis,output=TRUE,lineage.table=placenta.lineage.specific.analysis,members=5,orderbyfc=FALSE,fdr.cutoff=1) {
    temp.table <- 
        lineage.table[analysis.name==analysis &
                      p.value <= 0.05 &
                      fdr <= fdr.cutoff][order(p.value),
                                       list(human_name,p.value,fdr,fc),]
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <-
        literature.mining[,list(human_name,`Adverse Pregnancy Outcomes`)
                          ][temp.table]
    temp.table[,
               `Adv. Preg. Publications`:="0"
               ]
    temp.table[`Adverse Pregnancy Outcomes` >= 2,
               `Adv. Preg. Publications`:="≥ 2"
               ]
    temp.table[`Adverse Pregnancy Outcomes` >= 5,
               `Adv. Preg. Publications`:="≥ 5"
               ]
    temp.table[`Adverse Pregnancy Outcomes` >= 20,
               `Adv. Preg. Publications`:="≥ 20"
               ]
    temp.table[`Adverse Pregnancy Outcomes` == 1,
               `Adv. Preg. Publications`:="1"
               ]
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc,
                                            `Adv. Preg. Publications`
                                            )]
    setnames(temp.table,
             c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change",
               "Adv. Preg. Pub."
               )
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[`FDR`<=fdr.cutoff,][order(`p (t-test)`),]
        if (orderbyfc) {
            display.table <- temp.table[order(-abs(`Fold Change`)),]
        }
        display.table <- display.table[1:members,]

        display.table <-
            display.table[,`Adv. Preg. Pub.`:=
                               paste0("\\href{https://www.ncbi.nlm.nih.gov/pubmed/?term=%22",
                                      `Human Symbol`,
                                      "%22[All%20Fields]%20AND%20",
                                      literature.mining.terms,"}{",
                                      `Adv. Preg. Pub.`,"}"
                                      )]
        display.table <-
            display.table[,`Human Symbol`:=
                               paste0("\\href{http://www.ensembl.org/Homo_sapiens/Gene/Summary?g=",
                                      `Human Symbol`,"}{",
                                      "\\textit{",xtable.sanitize.text(`Human Symbol`),
                                      "}}")]
        display.table[,`Human Name`:=NULL]
        # setnames(display.table,"Human Symbol","\\multicolumn{1}{p{0.5in}}{Human Symbol}")
        setnames(display.table,"Fold Change","{$\\text{log}_2$ FC}")
        setnames(display.table,"FDR","{FDR}")
        setnames(display.table,"p (t-test)","{p (t-test)}")
        print(xtable(display.table,
                     display=c("s","s","e","e","f","s"),
                     align=c("c","p{0.5in}","S[table-format=1.2e2]","S[table-format=1.2e2]","S[table-format=2.2]","p{0.5in}")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=FALSE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

@@

\begin{table}[tbh]
  \footnotesize
<<human.lineage.specific,echo=FALSE,results="asis">>=
hum.all.lin.spec <-
    placenta.lineage.specific.table("hum.all",
                                    lineage.table=human.mouse.lineage.specific.analysis,
                                    members=40,
                                    orderbyfc=TRUE,
                                    fdr.cutoff=0.001
                                    )
setkey(hum.all.lin.spec,"Human Symbol")
@ 
\caption{\lineagetableexplanationtypeiiishort{\human}{all other
    analyzed species} Adv. Preg. Pub. is the approximate number of
  publications indexed by PubMed mentioning this gene and an adverse
  pregnancy outcome (pre-eclampsia, premature labor, gestational
  diabetes, intrauterine growth restriction, or premature birth) and
  links to the most current PubMed search results.}
\label{tab:human_lineage_specific}
\end{table}
